### 2022-04-10 Poster/poster.Rmd
poster for frontiers presentation

---
title: "Survey of Misuses of the Kolmogorov-Smirnov Test"
author:
  - name: Anthony Zeimbekakis
    affil: 1
  - name: Jun Yan
    affil: 1
affiliation:
  - num: 1
    address: Department of Statistics, University of Connecticut
column_numbers: 4
poster_height: "30in"
poster_width: "48in"
logoleft_name: "uconn-wordmark-single-white.png"
logoright_name: "uconn-wordmark-single-white.png"
primary_colour: "#000E2F"
secondary_colour: "#FFFFFF"
sectitle2_textcol: "#000E2F"
body_textsize: "30pt"
sectitle_textsize: "50pt"
column_margins: "0.2in"
output: 
  posterdown::posterdown_html:
    self_contained: false
bibliography: ../Manuscript/citations.bib
---

```{css, echo=FALSE}
#title {
  padding-top: 1.5cm;
  padding-bottom: 1.5cm;
}

.title_container {
  height: calc(38in * 0.13);
}

img {
  margin-top: 0cm;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1. Introduction

The Kolmogorov-Smirnov (K-S) test is one the most popular goodness-of-fit tests for 
comparing a sample with a hypothesized parametric distribution:

- Given a sample of n observations, let $S_{n}(x)$ be the empirical 
cumulative distribution and $F(x)$ be the population cumulative distribution
- The K-S statistic is defined by: \[d = max(\lvert F(x)-S_{n}(x) \rvert)\] 
- $H_0=$ The data follows a specified distribution $F(x)$
- $H_A=$ The data does not follow a specified distribution $F(x)$
- If the value of $d$ exceeds the test's corresponding critical value, 
the null hypothesis is rejected

The standard one-sample KS test applies to **independent, continuous data with a hypothesized distribution that 
is completely specified**, though it's often used incorrectly. We aim to survey misuses of the K-S test, 
demonstrate their consequences through simulation, and 
provide remedies as needed.

- All simulations unless specified are performed with the standard normal distribution $N(0,1)$

# 2. Fitted Parameters

## 2.1. Assumption:

- Hypothesized distribution $F(x)$ is completely specified

## 2.2. Misuse:

- Hypothesized distribution $F(x)$ is fitted with parameters from the sample

Ex: For the normal distribution, the test is performed where $F(x)$ ~ $N(\bar{X}, s^2)$ [@Lilliefors].

## 2.3. Remedy:

A remedy for this issue is to use bootstrap, detailed by [@Genest] and [@Babu]:

- $d_0$ is the observed K-S statistic
- Let $X_1^*,...,X_n^*$ be i.i.d random variables from $\hat{S}_n$, an estimator of the empirical distribution function $S$ based on the sample $X_1,...,X_n$
- Let $\hat{\theta}_n^* = \theta_n(X_1^*,...,X_n^*)$. $\theta$ represent the sample mean and variance of the bootstrap sample $X_1^*,...,X_n^*$
- New K-S statistic: $d = max(\lvert F(x)-S_{n}(x;\hat{\theta}_n^*) \rvert)$
- Count the number of bootstrap K-S statistics greater than or equal to $d_0$, and divide by the number of 
bootstrap samples to calculate p-value

## 2.4. Results:

<image width="100%" src="../Manuscript/hist_fitted.pdf"/>

- Naive plot was performed with $F(x)$ ~ $N(\bar{X}, s^2)$

- Parametric plot was performed with $\hat{S}_n = F(.;\hat{\theta}_n)$

- Babu plot was performed with $\hat{S}_n = S_n$ and a correction for bias from [@Babu]

With replicate tests, the distribution of p-values is $U(0,1)$. 
However since the assumption of a specified distribution is violated, the naive plot too 
often does not reject the null hypothesis. Bootstrap restores the power of the test.

# 3. Temporal Dependence

## 3.1 Assumption:

- The empirical distribution $S_{n}(x)$ contains independent data

## 3.2 Misues:

- The empirical distribution $S_{n}(x)$ contains dependent data

To demonstrate this, data is simulated from an $AR(1)$ process with $AR = (-3, 3)$.
There is an inverse relationship between negative and positive $AR$ values [@Durilleul], and the 
distributions are not uniform when temporal dependence is present.

<image width="100%" src="../Manuscript/hist_correlation.pdf"/>

## 3.3 Remedy:

- Use copulas combined with bootstrap, for two reasons
1. Ensure that the marginal distribution of our bootstrap data is the same as the fitted distribution
2. Introduce temporal dependence at the same level as the fitted distribution
The procedure used is the same as bootstrap, except the re-sampling comes from correlated data generated by
copulas

## 3.4 Results:

<image width="100%" src="../Manuscript/hist_copula_only.pdf"/>

- The copula remedy restores the $U(0,1)$ nature of the p-values

# 4. Fitted Parameters + Temporal Dependence

If both assumptions are violated, the copula fix works as well.

<image width="100%" src="../Manuscript/hist_copula.pdf"/>

- First plot shows the distribution of p-values when performing the copula remedy
- Second plot shows the distribution of p-values when performing basic parametric bootstrap,
which naively ignores dependence
- Third plot shows the violation of both assumptions,
where fitted parameters are used and dependence is not considered

# 5. Conclusion

- When the K-S test is used with violated assumptions, it no longer holds it's power
and corrections must be made

- In the case of fitted parameters, parametric and non-parametric bootstrap are recommended solutions

- When there is dependence in the data, performing bootstrap with copulas is recommended.

- When both assumptions are violated, meaning there is dependent data and parameters must be estimated,
perform bootstrap with copulas

- If the guess isn't far from the true distribution it may still be reasonably accurate for other distributions.

# 6. References