### 2022-04-28 Manuscript/manuscript.tex
various edits

\documentclass[12pt, letterpaper, titlepage]{article}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{setspace}

\usepackage[pagewise]{lineno}
%\linenumbers*[1]
% %% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
 \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
 \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
 \renewenvironment{#1}%
 {\linenomath\csname old#1\endcsname}%
 {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
 \patchAmsMathEnvironmentForLineno{#1}%
 \patchAmsMathEnvironmentForLineno{#1*}}%

\AtBeginDocument{%
 \patchBothAmsMathEnvironmentsForLineno{equation}%
 \patchBothAmsMathEnvironmentsForLineno{align}%
 \patchBothAmsMathEnvironmentsForLineno{flalign}%
 \patchBothAmsMathEnvironmentsForLineno{alignat}%
 \patchBothAmsMathEnvironmentsForLineno{gather}%
 \patchBothAmsMathEnvironmentsForLineno{multline}%
}

\newcommand{\jy}[1]{\textcolor{blue}{JY: #1}}

\title{On Misuses of the Kolmogorov--Smirnov Test for One Sample Goodness-of-Fit}

\author{Anthony Zeimbekakis\\
\href{mailto:anthony.zeimbekakis@uconn.edu}{\nolinkurl{anthony.zeimbekakis@uconn.edu}}\\
Department of Statistics, University of Connecticut\\\\
Thesis Advisor: Jun Yan\\
Honors Advisor: Elizabeth Schifano}
\date{}

\begin{document}
\maketitle

\doublespace

\begin{abstract}
The Kolmogorov--Smirnov (KS) test is one the most popular goodness-of-fit tests 
for comparing a sample with a hypothesized parametric distribution. Nevertheless, 
it has often been misused. The standard one-sample KS test applies to independent, 
continuous data with a hypothesized distribution that is completely specified. 
It is not uncommon, however, to see in the literature that it was applied to 
dependent, discrete, or rounded data, with hypothesized distributions containing 
estimated parameters. For example, it has been ``discovered'' multiple times 
that the test is too conservative when the parameters are estimated. We 
demonstrate misuses of the one-sample KS test in three scenarios through 
simulation studies:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
For each scenario, we provide remedies for practical applications.
\end{abstract}

\section{Introduction}\label{sec:intro}

The Kolmogorov-Smirnov (KS) statistic is one of the most popular goodness-of-fit 
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample from some continuous distribution. 
The null hypothesis $H_0$ is that the population distribution is $F$.
Let $F_n(\cdot)$ be the empirical cumulative distribution. The KS statistic is
\[
  D = \sup_x | F_{n}(x) - F(x) |.
\]
The distribution of $D$ under $H_0$ turned out to be independent of the
distribution $F$ and a table of critical values for $D$ has been constructed
\citep{Massey} for various sample sizes $n$ and significance 
levels $\alpha$. If the value of $d$ exceeds the test's corresponding critical 
value, the null hypothesis is rejected. The KS test is available in popular 
statistical software packages, such as function \texttt{ks.test} in R 
\citep{R, Marsaglia}.

The standard one-sample KS test applies to independent data with a continuous
hypothesized distribution that is completely specified.
\jy{this paragraph needs some reorganization.
  Flow: not applicable to discrete or rounded data (give references; which is
  not the focus of this paper);
  not applicable to fitted distributions;
  not applicable to serially dependent data}
\jy{break the big paragraph into smaller ones}
Often in literature, it has been "discovered" that the test loses its power when 
applied to dependent or discrete data with hypothesized distributions containing 
estimated parameters. DISCRETE SECTION HERE. This issue with discrete or rounded
is not the focus of this paper, though is a common misuse that should be 
noted.

The standard KS test is not applicable when the hypothesized distribution 
contains fitted parameters. \citet{Steinskog} demonstrates the change in power 
when using estimated parameters and stress caution in using the KS test in such 
ways. \citet{Lilliefors} shows that using the standard table when values of the 
mean and standard deviation are estimated obtains extremely conservative 
results. This is supported by \citet{Capasso}, which concludes that failing to 
re-estimate the parameters may lead to wrong, overly-conservative approximations 
to the distributions of goodness-of-fit test statistics based on the empirical 
distribution function. \citet{Capasso} also notes that the impact of this 
possible mistake may turn out to be dramatic and does not vanish as the sample 
size increases. Remedies are provided by \citet{Genest} and \citet{Babu} in the 
form of bootstrap. \citet{Genest} provides validity for using parametric 
bootstrap with various goodness-of-fit tests. \citet{Babu} details the bootstrap
procedure for goodness-of-fit tests and notes that both parametric and 
non-parametric procedures lead to correct asymptotic levels, however there 
is a correction required for the non-parametric case.

In the case of serially dependent data, \citet{Durilleul} demonstrates that the 
KS statistic is too liberal for medium-to-high positive autocorrelation values. 
\citet{Durilleul} also shows that for negative autocorrelation values, the 
behavior is asymmetrical with respect to positive values. For remedies, 
\citet{Weiss} provides a procedure that is applicable specifically for data 
modeled by the second-order auto-regressive (AR) process where the parameters 
are known. \citet{Lanzante} tests various strategies for dealing with temporal 
dependence and concludes that a test based on Monte-Carlo simulations performed 
the best. We propose a parametric bootstrap procedure involving copulas to 
account for dependence.

The contribution of this paper is a demonstration of misuses of the one-sample
KS test in three scenarios and their remedies in practice. The scenarios are 
where:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios. 
In each scenario, the misuse is performed and the impacts are shown. Then, a
remedy is detailed and performed alongside the misuse to show its positive 
effects. In order to set up the demonstrations, simulated data is used 
throughout. The remedies are also performed on a various families of 
distributions.

The rest of the paper is organized as follows. Section~\ref{sec:fitted}
investigates the scenario where the hypothesized distribution has unspecified
parameters. Both parametric and nonparametric bootstrap are available to fix the
issue. Section~\ref{sec:dependence} investigates the scenario where the data of 
the empirical distribution is serially dependent. Our bootstrap procedure 
employing copulas is a working solution. Section~\ref{sec:fittedwithdependence}
explores the case where a combination of the first two scenarios occurs. The 
copula procedure is available as a fix here as well. 
Section~\ref{sec:conclusion} concludes with a discussion.

\section{Fitted Parameters}\label{sec:fitted}

The null distribution of the KS statistic changes when the hypothesized
distribution contains fitted parameters. In this scenario, the null hypothesis
is $H_0$: the random sample $X_1, \ldots, X_n$ come from a continuous
distribution $F_{\theta}$ with unspecified parameter $\theta$.
Let $\hat\theta_n$ be an estimator of $\theta$, which could be, for example,
maximum likelihood estimator of moment estimator. The test statistic is
\[
  D = \sup_x | F_n(x) - F_{\hat\theta_n}(x) |.
\]
Since $F_{\hat\theta_n}$ is not the same as the true data generating
$F_\theta$, the null distribution of $D$ obtained in existing implementations in
software packages which assumes completely known $F_\theta$ no long applies
\citep{Steinskog}.

For illustration, \jy{give the details of the setting}
However, the procedure is sometimes performed where the population cumulative 
distribution $F(x)$ has parameters $\mu=\bar X$, the sample mean, and $\sigma^2=s^2$, 
the sample variance \citep{Lilliefors}. This case is demonstrated in Figure~\ref{fig:hist_fitted} 
under the Naive plot. The naive KS test was performed by estimating the parameters 
of $F(x)$ using the fitted distribution of the sample data. Since the data is generated 
from a standard normal distribution with seemingly all assumptions met, a uniform 
distribution of $U(0,1)$ is expected for the p-values. However this is under the 
assumption that the KS test holds its power, which it no longer does due to using 
fitted parameters. Therefore, there is notable deviation from the uniform distribution.

To fix the problem, parametric bootstrap can be used to approximate the null
distribution of the testing statistic. 
\begin{enumerate}
  \item 
    Draw a random sample $X_1^*,...,X_n^*$ from the fitted distribution $F_{\hat\theta_n}$
  \item 
    Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
  \item
    Obtain the empirical distribution function of the random sample $F_n^*$
  \item 
    Calculate bootstrap KS statistic
    \[
      D^* = \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) |.
    \]
  \item
    Repeat the previous steps a large number $B$ times and use the empirical 
    distribution of $D^*$ to approximate the null distribution of the observed 
    statistic.    
\end{enumerate}

Nonparametric bootstrap can also be used to approximate the null distribution 
of the testing statistic. The procedure is similar, however there is a 
correction for bias that is required \citep{Babu}.
\begin{enumerate}
  \item 
    Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
    with replacement
  \item 
    Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
  \item
    Obtain the empirical distribution function of the random sample $F_n^*$
  \item 
    Calculate bootstrap KS statistic
    \[
      D^* = \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) - B_n(x) |.
    \]
    where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known bias
    term \citep{Babu}
  \item
    Repeat the previous steps a large number $B$ times and use the empirical 
    distribution of $D^*$ to approximate the null distribution of the observed 
    statistic.
\end{enumerate}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_fitted}
  \caption{Implementing bootstrap remedies for fitted parameters misuse. \jy{make the caption self-contained.}}
  \label{fig:hist_fitted}
\end{figure}

Figure~\ref{fig:hist_fitted} displays the results of from our simulations. The 
procedure is replicated $1000$ times using the standard normal distribution with
sample size $n=100$. In each replicate test, $B = 1000$ bootstrap samples are 
obtained. The p-value can be calculated by counting the number of bootstrap KS 
statistics greater than or equal to the observed KS statistic, and then dividing 
by the number of bootstrap samples. It is clear from the figure that both
bootstrap processes correct for the problem of fitted parameters. The plots 
appear to be $U(0,1)$, which is expected of a full power test.

\section{Serially Dependent Data}\label{sec:dependence}

The KS test also displays issues in the case of dependence. As mentioned, an 
assumption of the test is that the data is independent. Unfortunately, real data
is often temporally or spatially dependent and the results of a goodness-of-fit 
test would be valuable. When the KS test is performed on dependent data it 
performs poorly. In this situation, the null hypothesis is $H_0$: the random 
sample $X_1, \ldots, X_n$ come from a continuous distribution $F_{\theta}$ with 
specified parameter $\theta$.
Let $\hat\theta_n$ be an estimator of $\theta$, which could be, for example,
maximum likelihood estimator of moment estimator. The test statistic is
\[
  D = \sup_x | F_n(x) - F_{\hat\theta_n}(x) |.
\]
However, since the data is dependent, the null hypothesis is wrong and must be 
corrected. This is demonstrated with a simulation in 
Figure~\ref{fig:hist_correlation}. Data is generated from a first-order 
autoregressive model (AR-1) with a standard normal distribution. The simulation
is done with the levels of the AR coefficient $\psi$, varying from $(-3,3)$. 
The results echo those of \citet{Durilleul} that the KS statistic is too liberal 
for positive autocorrelation values, and that the behavior is asymmetrical for 
negative values.

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_correlation}
  \caption{Results of varying AR levels of AR-1 simulated data.}
  \label{fig:hist_correlation}
\end{figure}

In order to correct this, we can employ a parametric bootstrap procedure which
assumes a working serial dependence structure through copulas. A copula is a
multivariate distribution with standard uniform marginal distributions, which
completely characterizes the dependence structure of a multivariate
distribution \citep{Copula, Hofert}. For simplicity, we assume a normal copula 
with an AR-1 structure to characterize the serial dependence of the observations. 
The AR-1 parameter of the normal copula is set to match the be the sample serial
Spearman's rho of the observed sample. The procedure is as follows.

\begin{enumerate}
  \item 
    Black box
    \jy{need to spell out the details; how to generate $X_t$ given $X_{t-1}$?}
  \item 
    Transform bootstrap sample to 
  \item
    Repeat the previous steps a large number $B$ times and use the empirical
    distribution of the $B$ test statistics as an approximate distribution of
    the null distribution of the observed statistic. 
\end{enumerate}

\jy{add a discussion about the method. The method is exact if the true
  dependence is indeed a normal copula with an ar1 structure. When it is not, it
  may still give a reasonable approximation that can be useful for practical purpose}

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_copula_only}
  \caption{Histograms demonstrating copula use to correct dependence.}
  \label{fig:hist_copula_only}
\end{figure}

Figure~\ref{fig:hist_copula_only} displays the results of the copula remedy for dependent data.
The data is generated from an AR-1 model with $\psi = .5$. The copula used is to model
dependence is the normal copula. $1000$ replicate tests were performed with
$B = 1000$ bootstrap samples per test.
\jy{present other scenarios too}
The naive test is performed by simply using the KS test without correcting for any 
potential dependence.

\section{Fitted Parameters and Serially Dependent Data}\label{sec:fittedwithdependence}

\jy{give a stronger motivation; last section is not practical at all because we
  often don't know the parameters; this section combines the last two}
The remedy using copulas is effective in the case where both assumptions are violated.
This is shown in Figure~\ref{fig:hist_copula}. The ``super naive" method is the resulting 
p-values when you naively use fitted parameters while ignoring dependence in the data.
The ``naive parametric" plot is the result of applying parametric bootstrap to correct
for fitted parameters while ignoring potential dependence. Of course, the "copula" 
plot applies the fix for both assumptions. The results show that copula remedy generates 
uniform p-values and restores the power of the KS test.

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_copula}
  \caption{Histograms demonstrating copula use to correct fitted parameters and dependence.}
  \label{fig:hist_copula}
\end{figure}

The copula fix is not a complete solution, because regardless of the true dependence in the data
we assume an AR(1) model by taking the lag-1 sample auto-spearman rho. However, we can 
show that as long as the AR(1) assumption is a close approximation, the correction still restores power to the test.
Figure~\ref{fig:hist_copula_ma1} is the procedure performed on data generated from 
an MA(1) model with $\theta = 0.5$. Figure~\ref{fig:hist_copula_arma} displays results with 
data generated from an ARMA(1,1) model with $\psi = 0.5$ and $\theta = 0.3$.

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_copula_ma1}
  \includegraphics{hist_copula_arma}
  \caption{MA-1 and ARMA(1,1)}
  \label{fig:hist_copula_ma1}
\end{figure}

It is also possible to apply our procedure to other distributions. All of the simulations
to this point have been done with standard normal data. The data in Figure~\ref{fig:hist_copula_gamma} 
is generated from $Gamma ~ (3,1)$ with a correlation coefficient of $\rho = 0.5$. Figure~\ref{fig:hist_copula_gev}
shows the procedure applied to data from $GEV ~ (0, 0.2, 1)$ with a correlation coefficient of $\rho = 0.5$.

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_copula_gamma}
  \includegraphics{hist_copula_gev}
  \caption{Gamma(3,1)}
  \label{fig:hist_copula_gamma}
\end{figure}

Where this remedy falls short is when the approximation of an AR-1 dependence 
structure is too inaccurate. An AR-2 model can be used to demonstrate this. For 
an AR-2 model with $\phi = (0.5, 0.3)$, Figure~\ref{fig:hist_copula_ar2} 
displays the results of applying the various remedies. While the remedies 
improve the plot of the p-values in comparison to the naive case, because the 
approximation is not close enough the correction does not work completely.

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_copula_ar2}
  \caption{AR-2}
  \label{fig:hist_copula_ar2}
\end{figure}

\section{Conclusion}\label{sec:conclusion}

The KS test has base assumptions that the hypothesized distribution is 
completely specified and the data is independent. When these assumptions are 
violated, the test is no longer accurate and remedies must be performed. In the 
case of fitted parameters, parametric and non-parametric bootstrap can restore 
the power of the test. A bias correction is required if the non-parametric form 
is used \citep{Babu}. In the case of dependent data, a procedure using bootstrap 
with copulas to model dependency shows positive results. When both assumptions 
are violated, that is the case where the data has dependence and parameters must 
be fitted, the copula remedy also shows positive results. The tests were 
performed on simulated data from a standard normal distribution, though also 
appear effective against other distributions such as gamma and gev. The copula 
remedy is not a complete solution. Regardless of the true dependence, we use an 
AR-1 model. Therefore, if the AR-1 model is a close approximation of the truth, 
the fix can work. However, in cases such as AR-2, the approximation too far and 
the fix does not completely restore power.

\bibliographystyle{chicago}
\bibliography{citations.bib}

\end{document}